---
title: "josh does a markdown"
output: html_document

---


In general, Rmarkdown separates sections via three ```. 

  * these sections are known as 'code chunks'
    + they organise your document for easier reading
        - you can quickly insert chunks by pressing **ctrl + alt + I** 
        - alternatively, you can add chunks with the add chunk button (green c with a plus)
        
For annotations that explain code, leave them out of code chunks.

# **Formatting**

There are many formatting options available in markdown including

- *italics*
- **bold**
- `code`
- you may also want to produce a~subscript~ or a^superscript^



In addition, you can make heading levels by adding # at the front of the line, these headings can also have in-line formatting;

# *hi* **there** 
## *hi* **there** 
### *hi* **there**

To produce the file you have been working on in Rmarkdown either click the knit button, or use **ctrl + shift + k**

For a detailed list of formatting please see this [link](https://pandoc.org/MANUAL.html#pandocs-markdown)


_______________________________________________________________

Rmarkdown provides an excellent method to integrate text into writing scripts, from simply annotating code for easier reading, up to producing documents and powerpoint presentations.

__________________________________________________________________


Images and figures are integrated straight into the document, and code can be run directly in the script to display the output while you work.

```{r fig.cap= "The Maunga Whau volcano", echo=FALSE}
image(volcano, col= hcl.colors(200, palette = 'viridis'))
```

___________________________________________________________________________
You can create, and alter dataframes much the same as you would in Rstudio

```{r}
rand_numb1 <- c(1, 3, 7, 5, 4, 7, 8, 3, 6)
speed <- c(10, 20, 30, 40, 50, 60, 70, 80, 90)
Rmkd_dataset <- data.frame(rand_numb1, speed)
head(Rmkd_dataset)
```

```{r}
summary(Rmkd_dataset)
```

```{r}
plot(Rmkd_dataset, type = "p", main = "speed at number= x")
```

_________________________________________________________________

# **Alternate code languages**



The knitr function also has functionaility in a variety of other porgramming languages, including;

- Python
- SQL
- Bash
- Rcpp
- Stan
- JavaScript
- CSS
  
All that is necessary is to change the **r** preface at the start of a code chunk

For example
#``` {bash}
#```

#```{python}
#```

etc.




The advantage of this is that it allows integration of the script across a multitude of different coding langauges.
This can be useful, for example if you have specific processes that requires programs not available in R

  - For example, in my case I used many program suites not available in R for data preparation and filtration.
    + At the time I had to maintain multiple different scripts across the programs to handle this, but had I used Rmarkdown it would have allowed me to contain it all within the single script.
    


    