---
title: "R markdown tutorial"
output: 
  html_document:
    toc: true
*output: 
  *html_notebook: default
  

---


## R markdown vs R markdown notebooks

The first thing I did was to change the output type from `html_document`, to `html_notebook: default` 
  
  *The benefit of this is that the notebook keeps a saved version of the document and incorporates additions, rather than having to re-render the entire document each time 
  
  *uses preview, rather than knit

In general, Rmarkdown separates sections via three ```. 



  * these sections are known as 'code chunks'
    * they organise your document for easier reading
        * you can quickly insert chunks by pressing **ctrl + alt + I** 
        * alternatively, you can add chunks with the add chunk button (green c with a plus)
        
For annotations that explain code, leave them out of code chunks.



## Formatting

There are many formatting options available in markdown including

- *italics*
- **bold**
- `code`
- you may also want to produce a~subscript~ or a^superscript^



In addition, you can make heading levels by adding # at the front of the line, these headings can also have in-line formatting;

# Header 1
## Header 2
### Header 3

To produce the file you have been working on in Rmarkdown either click the knit button, or use **ctrl + shift + k**

For a detailed list of formatting please see this [link](https://pandoc.org/MANUAL.html#pandocs-markdown)

The guide I used to create this can be found at: <https://www.dataquest.io/blog/r-markdown-guide-cheatsheet/>



# Bullet points and numbered lists

* List element 1 
* List element 2
    * List item 1
  
1. item 1
    * item 1a
1. item 2
1. item 3
    * item 3a
_______________________________________________________________

Rmarkdown provides an excellent method to integrate text into writing scripts, from simply annotating code for easier reading, up to producing documents and powerpoint presentations.

__________________________________________________________________


Images and figures are integrated straight into the document, and code can be run directly in the script to display the output while you work.


```{r volcano_Image, fig.cap= "The Maunga Whau volcano", echo=FALSE}
image(volcano, col= hcl.colors(200, palette = 'viridis'))

#note, to run this part of the script you need to have the package viridis installed in you R library- install.packages('viridis')
```



___________________________________________________________________________

You can create, and alter dataframes much the same as you would in Rstudio


```{r dataset_creation, echo=FALSE}
#creating dataset
rand_numb1 <- c(1, 3, 7, 5, 4, 7, 8, 3, 6)
speed <- c(10, 20, 30, 40, 50, 60, 70, 80, 90)
Rmkd_dataset <- data.frame(rand_numb1, speed)

#displaying dataset
head(Rmkd_dataset)

```

```{r data_summary, echo=FALSE}


summary(Rmkd_dataset)
```

```{r data_plot, echo=FALSE}
plot(Rmkd_dataset, type = "p", main = "speed at number= x")
```

_________________________________________________________________




# **Alternate code languages**



The knitr function also has functionaility in a variety of other porgramming languages, including;

- Python
- SQL
- Bash
- Rcpp
- Stan
- JavaScript
- CSS
  
All that is necessary is to change the **r** preface at the start of a code chunk

For example
#``` {bash}
#```

#```{python}
#```

etc.




The advantage of this is that it allows integration of the script across a multitude of different coding langauges.
This can be useful, for example if you have specific processes that requires programs not available in R

  - For example, in my case I used many program suites not available in R for data preparation and filtration.
    + At the time I had to maintain multiple different scripts across the programs to handle this, but had I used Rmarkdown it would have allowed me to contain it all within the single script.
    
    


    